Computing in the Field of Rationals

### 2012-06-19 rational_ext.tex
_

Andrzej Filinksi's masters thesis
\cite{Filinski:1989:DCI:648332.755574} suggested a remarkable symmetry
underlying computation in the \lcal. Filinski showed that values are
dual to continuations, functions are dual to delimited
continuations. Filinksi also shows a De Morgan style duality, by
treating value pairs as the dual of continuation sums and value sums
as dual to continuation pairs.

@@

\begin{proposition}[Correspondence]
  Evalution in the small-step evaulator corresponds to evaluation in
  the natural semantics.

@@

For a traced symmetric monoidal catgeory, the Int-construction of
Joyal, Street and Verity \cite{joyal1996traced} gives us a way of
constructing a compact closed category. This is the obvious first
thing to try for a duality in this setting.

@@

Provide a semantics with the nondeterminism monad and define an
\emph{eval}.

@@

Halmos \cite{halmos1958finite}, Abramsky and Coecke
\cite{abramsky2004categorical}. Despite the fact that this fragment
seem badly behaved (no logical reversibility) we have a close
connection with finite dimensional hilbert spaces.

The technical goal of this section is to show that the categorical
structure of finite dimensionl hilbert spaces with unitary maps and
inner product has the same categorical structure as {{PiEE*}}. Thus
constructions in {{PiEE*}} can be transported to FdHilb-unitary. If we
allow all linear maps, then {{+}} becomes a biproduct.  This has
connections to the Quantum Picturisms of Abramsky and Coecke and hence
to Qunatum Physics.

@@

There is much future work to be done in this line. The logical
consequence of this line of research is either (1) there is no way to
accomodate proabailities etc in this setting and this is all
meaningless or (2) constrcutions in {{PiEE*}} correspond to physical
phenomena.

>>>>

Andrzej Filinski's masters thesis
\cite{Filinski:1989:DCI:648332.755574} suggested a remarkable symmetry
underlying computation in the \lcal. Filinski showed that values are
dual to continuations, functions are dual to delimited
continuations. Filinksi also shows a De Morgan style duality, by
treating value pairs as the dual of continuation sums and value sums
as dual to continuation pairs.

@@

\begin{proposition}[Correspondence]
  Evaluation in the small-step evaluator corresponds to evaluation in
  the natural semantics.

@@

For a traced symmetric monoidal category, the Int-construction of
Joyal, Street and Verity \cite{joyal1996traced} gives us a way of
constructing a compact closed category. This is the obvious first
thing to try for a duality in this setting.

@@

Provide a semantics with the non-determinism monad and define an
\emph{eval}.

@@

Halmos \cite{halmos1958finite}, Abramsky and Coecke
\cite{abramsky2004categorical}. Despite the fact that this fragment
seem badly behaved (no logical reversibility) we have a close
connection with finite dimensional Hilbert spaces.

The technical goal of this section is to show that the categorical
structure of finite dimensional Hilbert spaces with unitary maps and
inner product has the same categorical structure as {{PiEE*}}. Thus
constructions in {{PiEE*}} can be transported to FdHilb-unitary. If we
allow all linear maps, then {{+}} becomes a biproduct.  This has
connections to the Quantum Picturisms of Abramsky and Coecke and hence
to Quantum Physics.

@@

There is much future work to be done in this line. The logical
consequence of this line of research is either (1) there is no way to
accommodate probabilities etc in this setting and this is all
meaningless or (2) constructions in {{PiEE*}} correspond to physical
phenomena.
